---
title: "Exercise set 1 -- Peter Tempfli"
output:
  html_document: default
  pdf_document: default
---
tempflip@gmail.com

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## 1.

The initial state
```{r}
vec1<-c(0,2,3,0,2,11,0,7,NA)
```



Use indexing to remove the NA.

```{r}
vec1 <- vec1[!is.na(vec1)]
vec1
```



Make a logical vector indicating the elements equal to zero

```{r}
myZeroPositions <- vec1==0
myZeroPositions
```



Use the logical vector to pick out the zero values and store them in a vector called ‘zeros’.
```{r}
zeros <- vec1[myZeroPositions]
zeros
```


Check how many zeros you have in vec1 by taking the length of the vector zeros (use function length()).
```{r}
length(zeros)
```

# 2.

```{r}
w <- c(109, 112, 115, 121, 128, 132, 135, 140, 148)
m <- c(120, 122, 124, 130, 136, 140, 143, 150, 155)
df <- data.frame(M = m, W = w)
row.names(df) <- seq(2003, 2011)
df <- t(df)

df
```

Construct a data frame from the table below, including the three variables W (average wage/h), YEAR (including the years for each observation) and Gender (including characters for Women/Men).

### Explanation:
I use the `melt` function in order to transform the data to long-format (= every observation has a separate row in the DF). Than I set manually the column names.

```{r}
library(reshape2)
each_observastion <- melt(df)
names(each_observastion) <- c('gender', 'year', 'wage')
each_observastion

setwd('~/prog/r')
write.csv(each_observastion, file='mydata.csv')

```

# 3.

``` {r}
library(openxlsx)
fr <- read.xlsx('~/Desktop/R/Freedman.xlsx')
fr$population <- as.numeric(fr$population)
fr$nonwhite <- as.numeric(fr$nonwhite)
fr$density <- as.numeric(fr$density)

summary(fr)
```

### Some basic visualization of the data

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(GGally)
ggpairs(fr[, -1])
```

# 4.
```{r}
library(car)
pr <- Prestige

```

Select a subset of the data for occupations with more than 50% women and call the subset ‘sub Prestige’.

```{r}
sub_Prestige <- pr[pr$women>50,]
sub_Prestige

```
Use the subset and compute the average prestige score.
```{r}
mean(sub_Prestige$prestige)
```

Now compute the average prestige score for occupations with less
than 50% women
```{r}
mean( subset(pr, women < 50)$prestige )
```

For this question use the complete Prestige data again. Make a for-loop to compute the average prestige score for the three dif- ferent types of occupations. Automatically store the three means in a vector. The loop should be general, i.e. even if the types of occupations were 100000 one should be able to use your loop.

### Explanation:

The function calculates the means of a property, grouped by another property. In the example above calculates the 'prestige' mean value, grouped by 'type'.

The group-by column must be a factor.


```{r}

mean_by_key <- function(df, valname, keyname) {

  counters <- rep(0, length(levels(df[,keyname]) ))
  sums <- rep(0, length(levels(df[,keyname]) ))

  for (i in seq(0:nrow(df))) {
    
    val <- df[i,valname]
    key <- df[i, keyname]
    counters[as.numeric(key)] <- counters[as.numeric(key)] + 1 
    sums[as.numeric(key)] <- sums[as.numeric(key)] + val
    
  }
  
  r <- data.frame(sums/counters)
  rownames(r) <- levels(df[,keyname])
  colnames(r) <- c(valname)
  return(r)
}

```

Demo
```{r}
pr <- Prestige
means_for_types <- mean_by_key(pr, 'prestige', 'type') ## column name and the grouped by are dynamic
means_for_types
```

We can check if our function output is correct, using the built-in `by` function.

```{r}
by(pr$prestige, pr$type, mean)
```

