---
title: "Exam_Peter_Tempfli"
author: "Peter Tempfli"
date: "4/2/2019"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)

```

## Sampling

I've created 3 functions:

* `hand` creates a hand (for example, [1,1,0,0,2]). Every number represents a color.
* `number_of_colors` counts the number of unique colors in a hand; the outcome of `hand` is passed to this functions
* `draws` creates N unique hands, and returns the number of hands which had at most 2 unique colors.

It is easy to see why `draws(N)/N` function converges to the probability of *at most 2 unique colors in a hand*. (so with a large enough N number we can get accurate enough results. Example: `draws(100000)/100000 = 0.14722`)


```{r}
hand <- function() {
  sample(c(rep(1,13), rep(2,13), rep(3,13), rep(4,14)), 5)  
}

number_of_colors <- function(h) {
  colors = c(sum(h==1), sum(h==2), sum(h==3), sum(h==4))
  return(sum(colors != 0))
}

draws <- function(n) {
  at_most_2 <- 0
  for (i in seq(n)) {
    if (number_of_colors(hand()) <= 2) at_most_2 = at_most_2 + 1
  }
  return (at_most_2)
}
```


```{r}
df <- data.frame(x=seq(500))
df$prob <- sapply(df$x, draws)
ggplot(data=df, aes(x=x, y=prob)) + geom_point()

```
